Parsing Postfix log files

### 2009-03-04 doc/logparser-architecture.tex
Wrote about pathological conditions. Dealt with XXX items.

Rules are responsible for categorising inputs: each rule should recognise
one and only input variant.  An input category with multiple input variants
should have multiple rules, one for each variant.  Rules will typically use
a regex when classifying inputs, but other approaches may prove useful for
some applications, e.g.\ comparing fixed strings to the input, or checking
the length of the input; for the remainder of this thesis it will be
assumed that a regex is used.  Each rule must specify at least the regex to
recognise inputs and the action to invoke when classification is
successful, but implementations are free to add any other attributes they
require; \sectionref{rule attributes} describes the attributes used in
\parsername{}, and some potential attributes will be discussed later in
\sectionref{attaching conditions to rules}.

Classifying inputs using the rules is simple: the first rule to recognise
the input determines the action that will be invoked: there is no
backtracking to try alternate rules; no attempt is made to pick a
\textit{best\/} rule.  \sectionref{attaching conditions to rules} contains
an example which requires that the rules are used in a specific order to
correctly parse the input, so a mechanism is needed to allow the author of
the rules to specify that ordering.  The framework should be written so
that it sorts the rules according to a priority attribute specified by the
rule author, allowing fine-grained control over the order that rules are
used in.  The priority attribute may be implemented as a number, or as a
range of values, e.g.\ low, medium, and high, or in a different fashion
entirely if it suits the implementation.  Rule ordering for efficiency is a
separate topic that is covered in \sectionref{rule ordering for
efficiency}; overlapping rules are discussed in \sectionref{overlapping
rules in architecture}.

@@

An example of how rule conditions can be useful is parsing C-style
comments, which start with \texttt{/*} and end with \texttt{*/}; the start
and end tokens can be on one line, or may have many lines between them.
\Tableref{Rules to parse C-style comments} shows the regexes, conditions,
and state changes of the four rules required to parse C-style comments.
These rules are somewhat simplified, e.g.\ rules one and two will recognise
the comment start token even if it is within a quoted string, which is
incorrect.  Rules one and two will only be used when the parser is parsing
code, not comments: rule one recognises a comment that is contained within
one line and leaves the parser's state unchanged; rule two recognises the
start of a command and changes the parser's state to parsing comments.
Rules three and four will be used when the parser is parsing a comment:
rule three recognises the end of a comment and switches parsing state back
to parsing code; rule four recognises a comment line without an end token
and keeps the parser's unchanged.  It is important that the rules are
applied in the order listed in \tableref{Rules to parse C-style comments};
\sectionref{rules in architecture} has explained how this is achieved, and
\sectionref{overlapping rules in architecture} discusses overlapping rules.

@@

Conditions can also check the value of variables that have been set by
either actions or rules; it will probably be easier to understand how a
variable's value will be used and changed if it is set by rules only rather
than by actions, because the chain of checking and setting can be followed
from rule to rule.  The downside to actions setting variables used in rule
conditions is action at a distance: to understand when a rule's condition
will be true requires understanding not just every rule but also every
action.  The framework probably does not need to support the same level of
complexity and flexibility when setting variables from rules (actions can
obviously use complex code if they are setting variables); however, if the
framework allows complex conditions, that code can probably be easily
extended to support complex variable assignments too.

@@

@@

XXX PATHOLOGICAL CONDITIONS\@.

\subsection{XXX TO BE DEALT WITH}

XXX DUNNO WHERE TO PUT THIS\@; IT SHOULD PROBABLY GO IN THE RULES SECTION,
BUT WHERE\@?

Rules are self-contained and can be understood in isolation, without
reference to other rules.  This is similar to a \acronym{CFG}; in
\acronym{CFG} terms the parser rules could be described as:

$\text{\textless{}input\textgreater{}} \mapsto{} \text{rule-1} |
\text{rule-2} | \text{rule-3} | \dots | \text{rule-n}$

Rules can not take into account past or future inputs when classifying the
current input, though conditions can affect which rules will be used to
classify each input.  The architecture is line oriented at present: there
is no facility for rules to consume more input or push unused input back
onto the input stream.  The current input can be manipulated by actions,
using the cascaded parsing feature.  This was not a deliberate design
decision, rather a consequence of the line oriented nature of Postfix log
files; more flexible approaches could be pursued.

\section{Conclusion}

>>>>

Rules are responsible for categorising inputs: each rule should recognise
one and only input variant.  An input category with multiple input variants
should have multiple rules, one for each variant.  Rules will typically use
a regex when classifying inputs, but other approaches may prove useful for
some applications, e.g.\ comparing fixed strings to the input, or checking
the length of the input; for the remainder of this thesis it will be
assumed that a regex is used.  Each rule must specify at least the regex to
recognise inputs and the action to invoke when classification is
successful, but implementations are free to add any other attributes they
require; \sectionref{rule attributes} describes the attributes used in
\parsername{}, and some potential attributes will be discussed later in
this section.

The architecture is line oriented at present: there is no facility for
rules to consume more input or push unused input back onto the input
stream.  This was not a deliberate design decision, rather a consequence of
the line oriented nature of Postfix log files; more flexible approaches
could be pursued.

Classifying inputs using the rules is simple: the first rule to recognise
the input determines the action that will be invoked; there is no
backtracking to try alternate rules; no attempt is made to pick a
\textit{best\/} rule.  \sectionref{attaching conditions to rules} contains
an example which requires that the rules are used in a specific order to
correctly parse the input, so a mechanism is needed to allow the author of
the rules to specify that ordering.  The framework should be written so
that it sorts the rules according to a priority attribute specified by the
rule author, allowing fine-grained control over the order that rules are
used in.  The priority attribute may be implemented as a number, or as a
range of values, e.g.\ low, medium, and high, or in a different fashion
entirely if it suits the implementation.  Rule ordering for efficiency is a
separate topic that is covered in \sectionref{rule ordering for
efficiency}; overlapping rules are discussed in \sectionref{overlapping
rules in architecture}.

In \acronym{CFG} terms the rules could be described as:

$\text{\textless{}input\textgreater{}}~\mapsto{}~\text{rule-1}~|~\text{rule-2}~|~\text{rule-3}~|~\dots~|~\text{rule-n}$

This is not entirely correct because the rules are not truly context free:
rule conditions (described in \sectionref{attaching conditions to rules})
restrict which rules will be used to recognise each input, imposing a
context of sorts.

@@

An example of how rule conditions can be useful is parsing C-style
comments, which start with \texttt{/*} and end with \texttt{*/}; the start
and end tokens can be on one line, or may have many lines between them.
\Tableref{Rules to parse C-style comments} shows the regexes, conditions,
and state changes of the four rules required to parse C-style comments.
These rules are somewhat simplified, e.g.\ rules one and two will recognise
the comment start token even if it is within a quoted string, which is
incorrect.  Rules one and two will only be used when the parser is parsing
code, not comments: rule one recognises a comment that is contained within
one line and leaves the parser's state unchanged; rule two recognises the
start of a command and changes the parser's state to parsing comments.
Rules three and four will be used when the parser is parsing a comment:
rule three recognises the end of a comment and switches parsing state back
to parsing code; rule four recognises a comment line without an end token
and keeps the parser's state unchanged.  It is important that the rules are
applied in the order listed in \tableref{Rules to parse C-style comments};
\sectionref{rules in architecture} has explained how this is achieved, and
\sectionref{overlapping rules in architecture} will discuss overlapping
rules.

@@

Conditions can also check the value of variables that have been set by
either actions or rules; it will be easier to understand how a variable's
value will be used and changed if it is set by rules only rather than by
actions, because the chain of checking and setting can be followed from
rule to rule.  The downside to actions setting variables used in rule
conditions is action at a distance: to understand when a rule's condition
will be true requires understanding not just every rule but also every
action.  The framework probably does not need to support the same level of
complexity and flexibility when setting variables from rules (actions can
obviously use complex code if they are setting variables); however, if the
framework allows complex conditions, that code can probably be easily
extended to support complex variable assignments too.

@@

Once conditions pass a certain level of complexity determining whether two
conditions overlap becomes intractable, because it requires so much
knowledge of other rules and possibly even actions.  E.g.\ given two rules
with conditions \verb!verbose == true! and \verb!silent == true!, it
appears that these rules cannot overlap, yet there is nothing to stop both
variables being set to true by one or more other rules or actions.  If
variables used in conditions can be set by actions, determining whether two
conditions overlap is impossible, because understanding how an action will
change the variables is impossible; the Church-Turing Thesis XXX NEED A BIB
ENTRY FOR CHURCH-TURING explains why.

@@

Conditions can vary in complexity from simple equality through to a
Turing-equivalent language, so enumerating pathological conditions is
difficult if not pointless.  Conditions that examine the input using
regexes can suffer from the same problems as rule regexes; conditions that
check variables in uncomplicated ways may exhibit unexpected or incorrect
behaviour, but are unlikely to exhibit pathological behaviour.  Deciding
whether a more complex condition's behaviour is pathological or simply a
bug is difficult and to some extent is a matter of opinion.  The author has
not experimented with more complex conditions, so this thesis will not
address this point any further.

\section{Summary}

### 2009-03-04 doc/logparser.tex
Dealt with some XXX items. Integrated some more content from the paper in the introduction section.
Improved the Framework section of the Rules chapter, and made some minor supporting changes elsewhere.
Dealt with some XXX items, and improved the pathological rules section.
Wrote about pathological conditions. Dealt with XXX items.

% Show the citation key (e.g. slct-paper) in the margin beside the
% reference.
%\usepackage[nomsgs]{refcheck}
% Show where references are used.
% Neither work.
%\usepackage{citeref}
%\usepackage{backref}
%\renewcommand{\refname}{Bibliography}
% Put each URL in the bibliography on a new line.
\newcommand{\urlprefix}[0]{\newline{}}
% New formatting commands.

@@

% Add last checked dates to all URLs
\newcommand{\urlLastChecked}[3][ ]{%
    \url{#2}#1(last checked #3)%
}
\newcommand{\daemonDocURL}[3][]{%
    \hfill{} \newline{} \url{#2}%
    \hfill{} #1 Last checked #3%
}
\newcommand{\urlLastCheckedNoParens}[3][ ]{%
    \url{#2}#1last checked #3%
}

@@

XXX POSSIBLY PUBLISH THE STATE OF THE ART REVIEW IN A JOURNAL\@?

@@

\renewcommand{\glossarypostamble}{\label{Glossary}}
\printglossary[style=nospacelist]{}
\renewcommand{\glossarypostamble}{\label{Acronyms}}
\printglossary[type=\acronymtype,style=eqlist]{}
\renewcommand{\glossarypostamble}{\label{Postfix Daemons}}
\printglossary[type=postfix,style=nospacelist]{}

>>>>

% Show the citation key (e.g.\ slct-paper) in the margin beside the
% reference.
%\usepackage[nomsgs]{refcheck}
% Show where references are used.
% Neither work.
%\usepackage{citeref}
%\usepackage{backref}
%\renewcommand{\refname}{Bibliography}

% Reduce the space between items in a list; this is useful when each item
% is a single line, because then the defaul spacing makes the list look
% very sparse.  This command needs to be the first thing in a list to take
% effect.
\newcommand{\squeezeitems}[0]{%
    \setlength{\itemsep}{0pt}%
}

% Put each URL in the bibliography on a new line.
\newcommand{\urlprefix}[0]{\newline{}}
% New formatting commands.

@@

% Add last checked dates to all URLs
\newcommand{\urlLastChecked}[3][ ]{%
    \url{#2}#1(last checked #3)%
}
\newcommand{\daemonDocURL}[3][]{%
    \hfill{} \newline{} \url{#2}%
    \hfill{} #1 Last checked #3%
}
% XXX CAN I MAKE THIS LOOK BETTER?  MAYBE USE [] LIKE \pageref{} DOES?
\newcommand{\urlLastCheckedNoParens}[3][ ]{%
    \url{#2}#1last checked #3%
}

@@

XXX ASK CARL ABOUT\@:

\begin{itemize}

    \item XXX WHERE SHOULD I USE URLS AND WHERE SHOULD THEY BE BIBLIOGRAPHY
        ENTRIES\@?

    \item XXX REFERENCE FOR TRANSDUCTION\@.

    \item XXX PREVIOUS RESEARCH IN OVERLAPPING RULES\@?

    \item XXX BORROW SPINNING WHEEL BOOK\@; NEED TO COMMENT ON THE
        COMPUTATIONAL COMPLEXITY OF FA UNION AND INVERSION\@; MAYBE COMMENT
        ON COMPUTATIONAL COMPLEXITY OF PDA UNION, INVERSION, AND
        INTERSECTION\@?

    \item XXX BIB ENTRY FOR CHURCH-TURING\@.

\end{itemize}

@@

% Add some glossary entries that should be present, but lack an appropriate
% place in the text to mark them.
\glsadd{queueid}
\renewcommand{\glossarypostamble}{\label{Glossary}}
\printglossary[style=nospacelist]{}
\renewcommand{\glossarypostamble}{\label{Acronyms}}
\printglossary[type=\acronymtype,style=eqlist]{}
\renewcommand{\glossarypostamble}{\label{Postfix Daemons}}
\printglossary[type=postfix,style=nospacelist]{}
